An Overview on Temporal Databases.\\ New Approaches and Proposals.

### 2012-03-01 trunk/book/author.tex
Ok, some modifications to the second chapter. Let me know what you think...
Last draft of the paper. This is the final draft.
The new form of the chapter...

\newcommand{\Pow}{\mathcal{P}}
\newcommand{\N}{\operatorname{N}}
\newcommand{\bool}{\operatorname*{\mathcal{B}}}
\newcommand{\Pos}{\operatorname{Pos}}
\newcommand{\Nec}{\operatorname{Nec}}
\newcommand{\T}{\mathcal{T}}
\newcommand{\C}{\mathcal{C}}
\newcommand{\Nat}{\mathbb{N}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\R}{\mathbb{R}}
\begin{document}

\title*{An Overview on Temporal Databases.\\
 New Approaches and Proposals.}
% Use \titlerunning{Short Title} for an abbreviated version of
% your contribution title if the original one is too long
\author{Jos\'e Pons \and Christophe Billiet \and Olga Pons \and Guy De Tr\'e}
% Use \authorrunning{Short Title} for an abbreviated version of
% your contribution title if the original one is too long
\institute{
Jos\'e Pons \at Department of Computer Science and Artificial Intelligence, Escuela T\'ecnica Superior de Ingenier\'ia Inform\'atica, Universidad de Granada,  C/Periodista Daniel Saucedo Aranda s/n, E-18071, Granada, Spain.
\email{jpons@decsai.ugr.es}
\and Christophe Billiet \at Department of Telecommunications and Information Processing, Ghent University, St.-Pietersnieuwstraat 41, B-9000 Gent, Belgium. \email{Christophe.Billiet@telin.ugent.be}
\and Olga Pons \at Department of Computer Science and Artificial Intelligence, Escuela T\'ecnica Superior de Ingenier\'ia Inform\'atica, Universidad de Granada,  C/Periodista Daniel Saucedo Aranda s/n, E-18071, Granada, Spain.
\email{opc@decsai.ugr.es}
\and Guy De Tr\'e \at Department of Telecommunications and Information Processing, Ghent University, St.-Pietersnieuwstraat 41, B-9000 Gent, Belgium. \email{Guy.DeTre@telin.ugent.be}
}
%
% Use the package "url.sty" to avoid
% problems with special characters
% used in your e-mail or web address
%
\maketitle

\vspace{-35pt}
%\abstract*{There exist some objects in the nature that have a time-variant (e.g. an employee's working life) or time-related (e.g. the transactions in a bank account) nature. Modelling these objects in a (relational) database is possible, but the time-related attributes have an impact on the consistency of the entire database. Therefore, some temporal database models have been proposed to deal with this. These models deal both with the problem of representation and querying. Time itself is a source of imprecision / vagueness / uncertainty since the measures of time are inherently imprecise. Humans beings manage the time with imprecision and vagueness. Nevertheless, the amount of uncertainty in the temporal indications are defined by socially well-known boundaries. Several proposals for dealing with imprecision in time have been done: some proposals consider the basis of the system the imprecision related to the conversion of diferent time points and intervals. While some other proposals consider the specification of the time as the source of imprecision: the time points are ill-known.}

\abstract*{
In reality, some objects or concepts have properties with a time-variant (e.g. an employee's job function) or time-related (e.g. financial transactions of a bank account) nature. Modelling these kinds of objects or concepts in a (relational) database schema is possible, but time-variant and time-related attributes have an impact on the consistency of the entire database. Therefore, some temporal database models have been proposed to deal with this. Time itself is a source of imprecision, vagueness and uncertainty, since existing time measuring devices are inherently imperfect. Accordingly, human beings manage time using temporal indications and temporal notions, which may contain imprecision, vagueness and uncertainty. However, the imperfection in human-used temporal indications is to some degree constrained by socially well-known boundaries. Several proposals for dealing with such imperfections exist. Some of these proposals consider the basis of the system to be the conversion of the specificity of temporal notions between used temporal expressions. Other proposals consider the temporal indications in the used temporal expressions to be the source of imperfection. In this chapter, an overview is given, concerning the basic concepts and issues related to the modelling of time as such or in (relational) database models and the imperfections that arise during or as a result of this modelling. Next to this, a novel and currently researched technique for handling some of these imperfections is presented.
}

\abstract{
In reality, some objects or concepts have properties with a time-variant (e.g. an employee's job function) or time-related (e.g. financial transactions of a bank account) nature. Modelling these kinds of objects or concepts in a (relational) database schema is possible, but time-variant and time-related attributes have an impact on the consistency of the entire database. Therefore, some temporal database models have been proposed to deal with this. Time itself is a source of imprecision, vagueness and uncertainty, since existing time measuring devices are inherently imperfect. Accordingly, human beings manage time using temporal indications and temporal notions, which may contain imprecision, vagueness and uncertainty. However, the imperfection in human-used temporal indications is to some degree constrained by socially well-known boundaries. Several proposals for dealing with such imperfections exist. Some of these proposals consider the basis of the system to be the conversion of the specificity of temporal notions between used temporal expressions. Other proposals consider the temporal indications in the used temporal expressions to be the source of imperfection. In this chapter, an overview is given, concerning the basic concepts and issues related to the modelling of time as such or in (relational) database models and the imperfections that arise during or as a result of this modelling. Next to this, a novel and currently researched technique for handling some of these imperfections is presented.
}

@@

\section{Time Modelling}
\label{sec:time-domain}
\input{sources/timeDomain.tex}
\def\JPicScale{1}
\section{Temporal Databases}
\label{sec:temporal-databases}
\input{sources/temporalDatabases.tex}
\def\JPicScale{1}
\section{Imprecision, Uncertainty and Vagueness in Temporal Databases}
\label{sec:imprecision-uncertainty}
\input{sources/fuzzyTemporalDatabases.tex}

@@

\newpage
\section*{Appendix: Introduction to fuzzy logic}
\label{app:fuzzy-logic}
\input{sources/fuzzylogic.tex}

>>>>

\newcommand{\Pow}{\mathcal{P}}
\newcommand{\N}{\operatorname{N}}
\newcommand{\bool}{\operatorname*{\mathcal{B}}}
\newcommand{\Pos}{\operatorname{Pos}}
\newcommand{\Nec}{\operatorname{Nec}}
\newcommand{\T}{\mathcal{T}}
\newcommand{\C}{\mathcal{C}}
\newcommand{\Nat}{\mathbb{N}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Z}{\mathbb{Z}}
\begin{document}

\title*{An Overview on Temporal Databases.\\
 New Approaches and Proposals.}
% Use \titlerunning{Short Title} for an abbreviated version of
% your contribution title if the original one is too long
\author{Jos\'e Pons \and Christophe Billiet \and Olga Pons \and Guy De Tr\'e}
% Use \authorrunning{Short Title} for an abbreviated version of
% your contribution title if the original one is too long
%\institute{
%Jos\'e Pons \at Department of Computer Science and Artificial Intelligence, Escuela T\'ecnica Superior de Ingenier\'ia Inform\'atica, Universidad de Granada,  C/Periodista Daniel Saucedo Aranda s/n, E-18071, Granada, Spain.
%\email{jpons@decsai.ugr.es}
%\and Christophe Billiet \at Department of Telecommunications and Information Processing, Ghent University, St.-Pietersnieuwstraat 41, B-9000 Gent, Belgium. \email{Christophe.Billiet@telin.ugent.be}
%\and Olga Pons \at Department of Computer Science and Artificial Intelligence, Escuela T\'ecnica Superior de Ingenier\'ia Inform\'atica, Universidad de Granada,  C/Periodista Daniel Saucedo Aranda s/n, E-18071, Granada, Spain.
%\email{opc@decsai.ugr.es}
%\and Guy De Tr\'e \at Department of Telecommunications and Information Processing, Ghent University, St.-Pietersnieuwstraat 41, B-9000 Gent, Belgium. \email{Guy.DeTre@telin.ugent.be}
%}

\institute{
Jos\'e Pons, Olga Pons \at Department of Computer Science and Artificial Intelligence, Escuela T\'ecnica Superior de Ingenier\'ia Inform\'atica, Universidad de Granada,  C/Periodista Daniel Saucedo Aranda s/n, E-18071, Granada, Spain.
\email{jpons, opc@decsai.ugr.es}
\and
Christophe Billiet, Guy De Tr\'e \at Department of Telecommunications and Information Processing, Ghent University, St.-Pietersnieuwstraat 41, B-9000 Gent, Belgium. \email{Christophe.Billiet, Guy.DeTre@telin.ugent.be}
}

%
% Use the package "url.sty" to avoid
% problems with special characters
% used in your e-mail or web address
%
\maketitle

\vspace{-70pt}
%\abstract*{There exist some objects in the nature that have a time-variant (e.g. an employee's working life) or time-related (e.g. the transactions in a bank account) nature. Modelling these objects in a (relational) database is possible, but the time-related attributes have an impact on the consistency of the entire database. Therefore, some temporal database models have been proposed to deal with this. These models deal both with the problem of representation and querying. Time itself is a source of imprecision / vagueness / uncertainty since the measures of time are inherently imprecise. Humans beings manage the time with imprecision and vagueness. Nevertheless, the amount of uncertainty in the temporal indications are defined by socially well-known boundaries. Several proposals for dealing with imprecision in time have been done: some proposals consider the basis of the system the imprecision related to the conversion of diferent time points and intervals. While some other proposals consider the specification of the time as the source of imprecision: the time points are ill-known.}

\abstract*{
In reality, some objects or concepts have properties with a time-variant or time-related nature. Modelling these kinds of objects or concepts in a (relational) database schema is possible, but time-variant and time-related attributes have an impact on the consistency of the entire database. Therefore, temporal database models have been proposed to deal with this. Time itself can be at the source of imprecision, vagueness and uncertainty, since existing time measuring devices are inherently imperfect. Accordingly, human beings manage time using temporal indications and temporal notions, which may contain imprecision, vagueness and uncertainty. However, the imperfection in human-used temporal indications is to some degree constrained by socially well-known boundaries. Several proposals for dealing with such imperfections when modelling temporal aspects exist. Some of these proposals consider the basis of the system to be the conversion of the specificity of temporal notions between used temporal expressions. Other proposals consider the temporal indications in the used temporal expressions to be the source of imperfection. In this chapter, an overview is given, concerning the basic concepts and issues related to the modelling of time as such or in (relational) database models and the imperfections that may arise during or as a result of this modelling. Next to this, a novel and currently researched technique for handling some of these imperfections is presented.
}

\abstract{
In reality, some objects or concepts have properties with a time-variant or time-related nature. Modelling these kinds of objects or concepts in a (relational) database schema is possible, but time-variant and time-related attributes have an impact on the consistency of the entire database. Therefore, temporal database models have been proposed to deal with this. Time itself can be at the source of imprecision, vagueness and uncertainty, since existing time measuring devices are inherently imperfect. Accordingly, human beings manage time using temporal indications and temporal notions, which may contain imprecision, vagueness and uncertainty. However, the imperfection in human-used temporal indications is to some degree constrained by socially well-known boundaries. Several proposals for dealing with such imperfections when modelling temporal aspects exist. Some of these proposals consider the basis of the system to be the conversion of the specificity of temporal notions between used temporal expressions. Other proposals consider the temporal indications in the used temporal expressions to be the source of imperfection. In this chapter, an overview is given, concerning the basic concepts and issues related to the modelling of time as such or in (relational) database models and the imperfections that may arise during or as a result of this modelling. Next to this, a novel and currently researched technique for handling some of these imperfections is presented.
}

@@

\section{Basic Concepts and Issues in Time Modelling}
\label{sec:time-domain}
\input{sources/timeDomain.tex}
\def\JPicScale{1}

\section{Data Imperfections in Time Modelling}
\label{sec:fuzzy-time-mod}
\input{sources/fuzzyTimeDomain.tex}

\section{Basic Concepts and Issues in Temporal Databases}
\label{sec:temporal-databases}
\input{sources/temporalDatabases.tex}
\def\JPicScale{1}

\section{Data Imperfections in Temporal Databases}
\label{sec:imprecision-uncertainty}
\input{sources/fuzzyTemporalDatabases.tex}

@@

%\newpage
%\section*{Appendix: Introduction to fuzzy logic}
%\label{app:fuzzy-logic}
%\input{sources/fuzzylogic.tex}

### 2012-03-01 trunk/book/sources/timeDomain.tex
The new form of the chapter...

%
% Time domain
%
Before considering the introduction of temporal modelling to information systems, it is necessary to define and explain some main concepts concerning temporal modelling and their corresponding terminology, to situate these and to discuss some properties and issues related to these concepts. In this section, several basic concepts and corresponding terminology will be defined, explained and situated. Most of these concepts are widely used in the community of temporal databases and their definitions have been agreed upon in the context of~\cite{Dyreson1994}. For these concepts, in the entire chapter, the contents of~\cite{Dyreson1994} are followed (and often cited).

In subsection \ref{subsec:basic-concepts}, some basic concepts of time modelling are presented and explained.

@@

\subsection{Basic concepts and properties}
\label{subsec:basic-concepts}
In information systems, time itself is usually perceived as a linear or cyclic concept~\cite{Jensen94thetsql2}. Therefore, a time domains modelling time is usually represented by a set with an imposed partial order. In general, two main types of time models can be discerned: a \emph{linear} model~\cite{benthem82} and a \emph{cyclic} model~\cite{lorentzos88}. In the linear model, a total order is imposed on the set and the progress of time is seen as a linear matter, while cyclic models are mainly used in the modelling of recurrent processes. It should be noted that the majority of proposals use linear time models.

@@

Orthogonal to the classification of underlying time models as linear or cyclic, they can be classified as \emph{discrete}, \emph{dense} or \emph{continuous} models. In a discrete model~\cite{Clifford:1985:AHR:971699.318922}, the notion exists that every instant has a unique successor and the set of instants is seen as a discrete one. Here, intuitively, the set of instants can be seen as isomorphic to the set of natural numbers $\Nat$. In a dense model, the notion exists that between any two instants always lies another, while the set of instants can still be seen as a discrete one. Here, intuitively, the set of instants can be seen as isomporphic to the set of rational numbers $\Q$ or the set of real numbers $\R$. In a continuous model, the notion also exists that between any two instants always lies another one, but unlike dense models, the set of instants is seen as continuous and there are no ``gaps'' between successive instants.

@@

The fact that chronons are actually time intervals has a particular effect on the representation of instants and time intervals. In a data model representing a time domain using chronons, an instant is of course represented as a chronon, but one single chronon may represent multiple instants. A time interval may be represented by a set of contiguous chronons, depending on the amount of time the time interval comprises.

@@

\textcolor{red}{CHECK}
Restrictions on time range may be performed on the base of the following two concepts~\cite{Jensen94thetsql2}:
\begin{itemize}
\item
\textbf{Time boundedness}: Time can be bounded orthogonally in the past and in the future.
\item
\textbf{Time distance}: As a metric, time has a distance function which presents four properties:
\begin{enumerate}
\item
Distance is non-negative.
\item
The distance between two different elements is not equal to zero.
\item
The distance from $\alpha$ to $\beta$ is the same as that between $\beta$ and $\alpha$.
\item
Triangle inequality: The distance from $\alpha$ to $\gamma$ is equal or greater than the distance from $\alpha$ to $\beta$ plus the distance from $\beta$ to $\gamma$.
\end{enumerate}
\end{itemize}
\textcolor{red}{END CHECK}

@@

Subsection \ref{subsec:granularity} gives a short survey on the concept of temporal granularities.

@@

As a granularity $G$ is an ordered set, each granule may be represented by an integer. In this representation, to keep track of the granularity a granule is an element of, the corresponding granularity is added in subscript:

@@

In a system, the granularity with the shortest granules is the \emph{chronons granularity}, which is denoted as $\bot$. It is the granularity of which the granules are chronons. 
%It is now possible to define a function $f$, called a \emph{mapping function}, that maps a given granule $i_G$, $i \in \Z$, from a given granularity $G$ to a set of corresponding chronons. The function definition should look like this:

\begin{svgraybox}
\begin{definition}
\label{def:mapping-function}
\textbf{Mapping function}~\cite{Lin97}\\ 
A mapping function $f$ is a function that translates from a given granule $i_G$, $i \in \Z$, in a given granularity $G$, to a set of corresponding chronons.\\

@@

\begin{align}
f & : G \rightarrow \Pow(\bot)\nonumber\\
  & : i_{G} \rightarrow \lbrace c_\bot\ |\ (c_\bot\text{ is contained in }i_{G}) \wedge (c_\bot \in \bot)\rbrace\nonumber
\end{align}
\end{definition}
\end{svgraybox}

Note that a mapping function $f$ always maps from a given granularity $G$ to the set of chronons $\bot$. Therefore, the output for a mapping function is a subset of $\Pow(\bot)$. \\
 A mapping function $f$ requires that the following properties hold~\cite{Lin97}:

@@

Mapping functions between granularities and the chronons granularity also allow comparing granularities with respect to the length of their granules. In this context, two important concepts can be discerned.
%A granularity $G$ is said to be \emph{\textbf{finner than}} a granularity $H$ when the set of chronons on which a given granule $i_G$ is mapped by a mapping function $f$ has less elements than the set of chronons on which a given granule $j_H$ is mapped by $f$.
%A granularity $\alpha$ is \emph{``coarser than''} granularity $\beta$ when the set of corresponding chronons for a given granule $i_\alpha$ is greater than the set of the corresponding chronons for a given granule $j_\beta$.

@@

It is also possible to describe the relation between different granularities. This is called a casting function:
\begin{svgraybox}
\begin{definition}
\textbf{Casting function}~\cite{Lin97}\\
\label{def:casting-fucntion}
A casting function $cast$ maps granules from a granularity $G$ to granules from a different granularity $H$.
\begin{equation}
\label{eq:cast-function}
\mbox{cast} \left( i_G,G,H \right) = j_H \nonumber
\end{equation}
\end{definition}
\end{svgraybox}

This notation indicates that the function $cast$ associates a granule $i_G$ in $G$ to a corresponding granule $j_H$ in $H$. Two kinds of granularity-to-granularity mappings can now be discerned: an \emph{upwards mapping} is a mapping from a granularity $G$ to a coarser granularity $H$, whereas a downwards mapping is a mapping from a granularity $K$ to a finner granularity $L$. Different granularity-to-granularity mappings between several granularities can be represented using a granularity graph, which is a directed graph indicating the mapping conversions.

 Mappings between two granularities may be classified as \emph{regular mappings}, \emph{irregular mappings} or \emph{congruent mappings}.
\begin{itemize}
\item
\emph{Regular mapping}: A regular mapping is a granularity-to-granularity mapping where the mapping function between the two granularities is calculated by means of multiplications and/or divisions and (maybe) an anchor adjustment. E.g. the mapping between hours and minutes is calculated using a multiplication by $60$. 
\item
\emph{Irregular mapping}: An irregular mapping is a granularity-to-granularity mapping where the mapping function can not be calculated by means of multiplications and/or divisions. E.g. the mapping between months and days is dependent on the exact month or day. 
\item
\emph{Congruent mapping}: A congruent mapping is a granularity-to-granularity mapping where the two granularities in the mapping have the same granules but a different anchor. E.g. the mapping between the days (Gregorian calendar days) and the academic days. 
\end{itemize}

\begin{example}
Consider a systems that models both Gregorian Calendar dates as well as Academic dates. The chronons granularity in this system are milliseconds. Figure \ref{fig:granularity-graph-example} shows the complete granularity graph example. The transitions between Chronons and Seconds and between Seconds and Minutes and so on is called a \emph{regular mapping}. The transition between Days and Months is not defined by a multiplication or division, this is called a \emph{irregular mapping}. In the graph, these \emph{irregular mappings} are noted with a bold arrow. Finally, the transition between a Gregorian Calendar Day and an Academic Day is called a \emph{Congruent mapping}: two granularities with the same granules but with different anchors. Both concern the same days, but the academic year starts on October 1st, whereas the Gregorian calendar year starts on January 1st. Congruent mappings are visualised as straight lines without arrow heads.

>>>>

%
% Time domain
%
Before considering the introduction of temporal modelling to information systems, it is necessary to define and explain some main concepts concerning temporal modelling and their corresponding terminology, to situate these and to discuss some properties and issues related to these concepts. In this section, several basic concepts and their corresponding terminology will be defined, explained and situated. Most of these concepts are widely used in the community of temporal databases and their definitions have been agreed upon in the context of~\cite{Dyreson1994}. For these concepts, in the entire chapter, the contents of~\cite{Dyreson1994} are followed (and often cited).

%In subsection \ref{subsec:basic-concepts}, some basic concepts of time modelling are presented and explained.

@@

\subsection{Basic Concepts and Properties}
\label{subsec:basic-concepts}
In information systems, time itself is usually perceived as a linear or cyclic concept~\cite{Jensen94thetsql2}. Therefore, a time domain modelling time is usually represented by a set with an imposed partial order. In general, two main types of time models can be discerned: a \emph{linear} model~\cite{benthem82} and a \emph{cyclic} model~\cite{lorentzos88}. In the linear model, a total order is imposed on the set and the progress of time is seen as a linear matter, while cyclic models are mainly used in the modelling of recurrent processes. It should be noted that the majority of proposals use linear time models.

@@

Orthogonal to the classification of underlying time models as linear or cyclic, they can be classified as \emph{discrete}, \emph{dense} or \emph{continuous} models\cite{Dyreson1994}. In a discrete model~\cite{Clifford:1985:AHR:971699.318922}, the notion exists that every instant has a unique successor and the set of instants is seen as a discrete one. Here, intuitively, the set of instants can be seen as isomorphic to the set of natural numbers $\Nat$. In a dense model, the notion exists that between any two instants always lies another. Here, intuitively, the set of instants can be seen as isomporphic to the set of rational numbers $\Q$ (when the set of instants is a discrete one) or the set of real numbers $\R$ (when the set of instants is a continuous one). In a continuous model, the notion also exists that between any two instants always lies another one, but the set of instants is always seen as continuous and there are no ``gaps'' between successive instants.

@@

The fact that chronons are actually time intervals has a particular effect on the representation of instants and time intervals. In a data model representing a time domain using chronons, an instant is of course represented by a chronon. A time interval may be represented by a set of contiguous chronons, depending on the amount of time the time interval comprises.

@@

%\textcolor{red}{CHECK}
Restrictions on time range may exist, as time may be bounded orthogonally in the past and in the future~\cite{Jensen94thetsql2}.
%\begin{itemize}
%\item
%\textbf{Time boundedness}: Time can be bounded orthogonally in the past and in the future.
%\item
%\textbf{Time distance}: As a metric, time has a distance function which presents four properties:
%\begin{enumerate}
%\item
%Distance is non-negative.
%\item
%The distance between two different elements is not equal to zero.
%\item
%The distance from $\alpha$ to $\beta$ is the same as that between $\beta$ and $\alpha$.
%\item
%Triangle inequality: The distance from $\alpha$ to $\gamma$ is equal or greater than the distance from $\alpha$ to $\beta$ plus the distance from $\beta$ to $\gamma$.
%\end{enumerate}
%\end{itemize}
%\textcolor{red}{END CHECK}

@@

%Subsection \ref{subsec:granularity} gives a short survey on the concept of temporal granularities.

@@

As a granularity $G$ is an ordered set, each granule may be represented by an integer. In this representation, to keep track of the granularity a granule is an element of, the corresponding granularity name is added in subscript:

@@

In a system, the granularity with the shortest granules is the \emph{chronons granularity}, which is denoted as `$\bot$'. It is the granularity of which the granules are chronons. 
%It is now possible to define a function $f$, called a \emph{mapping function}, that maps a given granule $i_G$, $i \in \Z$, from a given granularity $G$ to a set of corresponding chronons. The function definition should look like this:

\begin{svgraybox}
\vspace{-10pt}
\begin{definition}
\label{def:mapping-function}
\textbf{Mapping function}~\cite{Lin97}\\ 
A mapping function $f$ is a function that maps a given granule $i_G$, $i \in \Z$, in a given granularity $G$, to a set of corresponding chronons:\\

@@

\begin{align}
f & : G \rightarrow \Pow(\bot)\nonumber\\
  & : i_{G} \rightarrow \lbrace c_\bot\ |\ (c_\bot\text{ is contained in }i_{G}) \wedge (c_\bot \in \bot)\rbrace\nonumber
\end{align}
\end{definition}
\vspace{-10pt}
\end{svgraybox}

Note that a mapping function $f$ always maps from a given granularity $G$ to the powerset of the set of chronons $\bot$. Therefore, the output for a mapping function is an element of $\Pow(\bot)$ and thus a subset of $\bot$. \\
 A mapping function $f$ requires that the following properties hold~\cite{Lin97}:

@@

The existence of mapping functions between granularities and the chronons granularity also allows comparing granularities with respect to the length of their granules. In this context, two important concepts can be discerned.
%A granularity $G$ is said to be \emph{\textbf{finner than}} a granularity $H$ when the set of chronons on which a given granule $i_G$ is mapped by a mapping function $f$ has less elements than the set of chronons on which a given granule $j_H$ is mapped by $f$.
%A granularity $\alpha$ is \emph{``coarser than''} granularity $\beta$ when the set of corresponding chronons for a given granule $i_\alpha$ is greater than the set of the corresponding chronons for a given granule $j_\beta$.

@@

It is also possible to describe the relation between different granularities. This is called a casting function:
\begin{svgraybox}
\begin{definition}
\textbf{Casting function}~\cite{Lin97}\\
\label{def:casting-fucntion}
Consider two different granularities $G$ and $H$. A granularity-to-granularity \emph{\textbf{casting function}} $cast$ is then a function mapping granules from $G$ to granules from $H$: 
\begin{align}
\label{eq:cast-function}
%\mbox{cast} \left( i_G,G,H \right) = j_H \nonumber
cast & : G \times \mathbb{G} \times \mathbb{G} \rightarrow H \nonumber\\
		 & : (i_{G}, G, H) \rightarrow j_{H} \nonumber
\end{align}
where $i_{G} \in G$ and $j_{H} \in H$ and where $\mathbb{G}$ denotes the set of all granularities.
\end{definition}
\end{svgraybox}

Thus, the function $cast$ associates a granule $i_G$ in $G$ to a corresponding granule $j_H$ in $H$. Two kinds of granularity-to-granularity mappings can now be discerned: an \emph{upwards mapping} is a mapping from a granularity $G$ to a coarser granularity $H$, whereas a downwards mapping is a mapping from a granularity $K$ to a finner granularity $L$. Different granularity-to-granularity mappings between several granularities can be represented using a granularity graph, which is a directed graph indicating the mapping conversions.

Mappings between two granularities may be classified as \emph{regular mappings}, \emph{irregular mappings} or \emph{congruent mappings}~\cite{Lin97},~\cite{DyresonSnodgrass1994}.
\begin{itemize}
\item
\emph{Regular mapping}: A regular mapping is a granularity-to-granularity mapping where the mapping function value is calculated by means of multiplications and/or divisions and (maybe) an anchor adjustment. E.g. the mapping value of the mapping between hours and minutes is calculated using a multiplication by $60$. 
\item
\emph{Irregular mapping}: An irregular mapping is a granularity-to-granularity mapping where the mapping function value can not be calculated by means of multiplications and/or divisions. E.g. the mapping value of the mapping between months and days is dependent on the exact month or day. 
\item
\emph{Congruent mapping}: A congruent mapping is a granularity-to-granularity mapping where the two granularities involved in the mapping have the same granules but a different anchor. E.g. the mapping between the days (Gregorian calendar days) and the academic days is a congruent mapping.
\end{itemize}

The above is illustrated in the following example.

\begin{example}
Consider a system that models both Gregorian Calendar dates as well as Academic dates. The chronons granularity in this system contains milliseconds. Figure \ref{fig:granularity-graph-example} shows the complete granularity graph corresponding with this example. The transitions between Chronons and Seconds and between Seconds and Minutes and so on is called a \emph{regular mapping}. The transition between Days and Months is not defined by a multiplication or division, this is called a \emph{irregular mapping}. In the graph, these \emph{irregular mappings} are noted with a bold arrow. Finally, the transition between a Gregorian Calendar Day and an Academic Day is called a \emph{Congruent mapping}: two granularities with the same granules but with different anchors. Both concern the same days, but the academic year starts on October 1st, whereas the Gregorian calendar year starts on January 1st. Congruent mappings are visualised as straight lines without arrow heads.
